---
title: "A/B Testing a Call to Action"
subtitle: "Simulating Key Ideas from Classical Frequentist Statistics"
author: "Ria Tang"
date: April 14, 2026
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    theme: flatly
    fig-width: 8
    fig-height: 5
execute:
  warning: false
  message: false
---

```{r setup, echo=false}
library(tidyverse)
library(gt)
set.seed(42)
```

## Introduction

Imagine you run a website with a newsletter, and you want as many visitors as possible to sign up. On your landing page, you have a single "call to action" (CTA) — a short line of text that prompts visitors to subscribe. You're currently testing two versions:

- **CTA A:** *"Sign up for our newsletter here!"*
- **CTA B:** *"Stay up to date by signing up!"*

Which one actually works better? Your gut feeling isn't enough — you need data. That's where an **A/B test** comes in.

In an A/B test, each visitor to your site is randomly assigned to see one of the two CTAs. By randomly splitting visitors into two groups, we ensure that any difference in sign-up rates between the groups can be attributed to the CTA itself, not to some other difference between the visitors. After collecting enough data, we compare the sign-up rates across the two groups and use statistical tools to decide whether any observed difference is real or just due to random chance.

In this post, we'll walk through the key ideas from classical frequentist statistics that underpin a proper A/B test — using simulation to make each concept concrete and intuitive.

---

## The A/B Test as a Statistical Problem

To analyze our A/B test rigorously, we need to frame it as a formal statistical problem.

Each visitor to our site either signs up (outcome = 1) or doesn't (outcome = 0). This binary outcome is naturally modeled as a draw from a **Bernoulli distribution**: a coin flip with some probability $\pi$ of landing heads (signing up). Because the two CTAs might influence visitor behavior differently, we allow the sign-up probability to differ between groups:

- Visitors who see **CTA A** have sign-up probability $\pi_A$
- Visitors who see **CTA B** have sign-up probability $\pi_B$

The quantity we actually care about is the **difference in sign-up rates**:

$$\theta = \pi_A - \pi_B$$

If $\theta > 0$, CTA A is better. If $\theta < 0$, CTA B is better. If $\theta = 0$, neither outperforms the other.

Of course, we never observe $\pi_A$ or $\pi_B$ directly — we only see the outcomes for the visitors in our experiment. So we estimate $\theta$ using the **difference in sample proportions**:

$$\hat\theta = \bar{X}_A - \bar{X}_B = \hat\pi_A - \hat\pi_B$$

where $\bar{X}_A$ is the fraction of CTA A visitors who signed up, and $\bar{X}_B$ is the same for CTA B. This estimator $\hat\theta$ is the central object of our analysis. The rest of this post explores its statistical properties through simulation.

---

## Simulating Data

In the real world, we would never know the true values of $\pi_A$ and $\pi_B$ — that's precisely why we run the experiment. But for the purposes of this exercise, we get to play God: we'll set the true parameters ourselves, run our statistical methods, and verify that they behave as theory predicts.

We'll assume:

$$\pi_A = 0.22, \quad \pi_B = 0.18, \quad \theta = \pi_A - \pi_B = 0.04$$

So CTA A is genuinely better by 4 percentage points. We simulate 1,000 visitors per group — 2,000 observations total.

```{r simulate-data}
set.seed(42)

pi_A <- 0.22
pi_B <- 0.18
n    <- 1000

draws_A <- rbinom(n, 1, pi_A)
draws_B <- rbinom(n, 1, pi_B)

cat("Sample proportion for CTA A:", round(mean(draws_A), 4), "\n")
cat("Sample proportion for CTA B:", round(mean(draws_B), 4), "\n")
cat("Estimated difference (theta-hat):", round(mean(draws_A) - mean(draws_B), 4), "\n")
```

Our sample proportions are very close to the true values of 0.22 and 0.18 — a reassuring sign, but why exactly can we trust them? That's what the next section is about.

---

## The Law of Large Numbers

The **Law of Large Numbers (LLN)** tells us that as the sample size grows, the sample mean converges to the true population mean. In our setting, this means $\hat\pi_A \to \pi_A$ and $\hat\pi_B \to \pi_B$ as we collect more data — and therefore $\hat\theta \to \theta$.

The LLN is the foundational reason we trust our estimator: given enough data, it will get arbitrarily close to the truth. But how much data is "enough"? The plot below makes this tangible.

We compute the element-wise difference between our CTA A and CTA B draws, then track the **running (cumulative) average** of those differences as we add one observation at a time.

```{r lln-plot, echo=true}
diffs       <- draws_A - draws_B
running_avg <- cumsum(diffs) / seq_along(diffs)

lln_df <- tibble(
  n   = seq_along(running_avg),
  avg = running_avg
)
```

```{r lln-chart, echo=false}
ggplot(lln_df, aes(x = n, y = avg)) +
  geom_line(color = "#4e79a7", linewidth = 0.6, alpha = 0.85) +
  geom_hline(yintercept = 0.04, linetype = "dashed", color = "#e15759", linewidth = 0.8) +
  annotate("text", x = 850, y = 0.046, label = "True θ = 0.04",
           color = "#e15759", size = 3.5, fontface = "italic") +
  scale_y_continuous(limits = c(-0.15, 0.25)) +
  labs(
    title    = "Law of Large Numbers: Running Average of CTA A − CTA B",
    subtitle = "The estimate stabilizes around the true difference as sample size grows",
    x        = "Number of Observations",
    y        = "Cumulative Average Difference"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title    = element_text(face = "bold"),
    plot.subtitle = element_text(color = "grey40")
  )
```

With only a handful of observations, the running average is all over the place — swinging from strongly positive to slightly negative. But as the sample size grows, the noise dies down and the estimate settles close to the true value of 0.04 (the red dashed line). This is the LLN at work: more data means a more reliable estimate.

The practical takeaway for A/B testing is that **small samples can be deeply misleading**. If you had stopped at 50 visitors per group, you might have drawn completely wrong conclusions. Size matters.

---

## Bootstrap Standard Errors

Knowing that our estimator $\hat\theta$ is close to the truth on average is useful, but it raises a follow-up question: *how close?* A point estimate without a measure of uncertainty is incomplete. We need a **confidence interval**.

One way to quantify the uncertainty in $\hat\theta$ is the **bootstrap**. The idea is elegantly simple: instead of deriving a formula for the variability of our estimator, we use the data itself to simulate what would happen if we ran the experiment many times.

Here's the procedure:

1. From our 1,000 CTA A observations and 1,000 CTA B observations, draw a **resample of size 1,000 with replacement** from each group.
2. Compute $\hat\theta^* = \bar{X}_A^* - \bar{X}_B^*$ on the resample.
3. Repeat 1,000 times. The standard deviation of the 1,000 bootstrapped $\hat\theta^*$ values is our **bootstrapped standard error**.

```{r bootstrap, echo=true}
B <- 1000
boot_estimates <- numeric(B)

for (b in 1:B) {
  resample_A       <- sample(draws_A, size = n, replace = TRUE)
  resample_B       <- sample(draws_B, size = n, replace = TRUE)
  boot_estimates[b] <- mean(resample_A) - mean(resample_B)
}

se_boot <- sd(boot_estimates)

# Analytical SE for difference in two independent Bernoulli proportions
pi_hat_A <- mean(draws_A)
pi_hat_B <- mean(draws_B)
se_analytical <- sqrt(pi_hat_A * (1 - pi_hat_A) / n +
                      pi_hat_B * (1 - pi_hat_B) / n)

theta_hat <- pi_hat_A - pi_hat_B
ci_lower  <- theta_hat - 1.96 * se_boot
ci_upper  <- theta_hat + 1.96 * se_boot
```

```{r se-table, echo=false}
tibble(
  Quantity = c(
    "Point Estimate (θ̂)",
    "Bootstrapped Standard Error",
    "Analytical Standard Error",
    "95% Confidence Interval"
  ),
  Value = c(
    round(theta_hat, 4),
    round(se_boot, 4),
    round(se_analytical, 4),
    NA
  ),
  Notes = c(
    paste0("True θ = 0.04"),
    "SD of 1,000 bootstrap resamples",
    "Closed-form formula",
    paste0("[", round(ci_lower, 4), ", ", round(ci_upper, 4), "]")
  )
) |>
  gt() |>
  fmt_missing(columns = Value, missing_text = "—") |>
  tab_header(
    title    = "Bootstrap vs. Analytical Standard Errors",
    subtitle = "Both methods agree closely"
  ) |>
  cols_label(Quantity = "Quantity", Value = "Value", Notes = "Notes") |>
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_options(table.width = pct(80))
```

The bootstrapped and analytical standard errors are very close to each other — this is expected and reassuring. The bootstrap "discovers" the right answer without needing a formula, purely from the data.

Our **95% confidence interval** says: given our data, we estimate that CTA A's sign-up rate exceeds CTA B's by somewhere between roughly `r round(ci_lower*100,1)`% and `r round(ci_upper*100,1)`% percentage points. Since the entire interval sits above zero, this is strong evidence that CTA A genuinely outperforms CTA B.

---

## The Central Limit Theorem

The bootstrap gave us a practical tool for computing standard errors, but there's a deeper theoretical result underpinning our ability to do inference: the **Central Limit Theorem (CLT)**.

The CLT states that regardless of the shape of the underlying population distribution, the sampling distribution of the sample mean becomes approximately **Normal** as the sample size grows. In our case, even though individual outcomes are Bernoulli (binary 0/1), the distribution of $\hat\theta = \bar{X}_A - \bar{X}_B$ will look like a bell curve for large enough $n$.

To see this in action, we simulate 1,000 experiments at four different sample sizes — $n \in \{25, 50, 100, 500\}$ — and plot the distribution of $\hat\theta$ each time.

```{r clt-sim, echo=true}
sample_sizes <- c(25, 50, 100, 500)
n_sim        <- 1000

clt_results <- map_dfr(sample_sizes, function(n_s) {
  theta_hats <- replicate(n_sim, {
    mean(rbinom(n_s, 1, pi_A)) - mean(rbinom(n_s, 1, pi_B))
  })
  tibble(n = n_s, theta_hat = theta_hats)
}) |>
  mutate(n_label = paste0("n = ", n),
         n_label = factor(n_label, levels = paste0("n = ", sample_sizes)))
```

```{r clt-plot, echo=false}
ggplot(clt_results, aes(x = theta_hat)) +
  geom_histogram(aes(y = after_stat(density)),
                 bins = 35, fill = "#4e79a7", color = "white", alpha = 0.85) +
  geom_vline(xintercept = 0.04, linetype = "dashed",
             color = "#e15759", linewidth = 0.8) +
  facet_wrap(~n_label, scales = "fixed", nrow = 2) +
  labs(
    title    = "Central Limit Theorem: Sampling Distribution of θ̂ at Different Sample Sizes",
    subtitle = "Red dashed line marks the true θ = 0.04",
    x        = "θ̂ = X̄_A − X̄_B",
    y        = "Density"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title    = element_text(face = "bold"),
    plot.subtitle = element_text(color = "grey40"),
    strip.text    = element_text(face = "bold", size = 12)
  )
```

The progression is striking. At $n = 25$, the histogram is jagged and irregular — the distribution of $\hat\theta$ doesn't look remotely Normal. By $n = 100$, a clear bell shape is emerging. At $n = 500$, the distribution is smooth and symmetric around the true value of 0.04.

This matters enormously for practice: the statistical tests we run in the next section rely on the Normal approximation. The CLT tells us that approximation is valid for large samples — which is why serious A/B tests require hundreds of observations per group, not dozens.

---

## Hypothesis Testing

Now that we've seen the CLT in action, we can use it to run a formal **hypothesis test**.

We set up two competing hypotheses:

- **Null hypothesis** $H_0$: $\theta = 0$ (the two CTAs produce identical sign-up rates)
- **Alternative hypothesis** $H_1$: $\theta \neq 0$ (the CTAs differ)

The goal is to assess whether the data provide enough evidence to reject $H_0$. We do this by asking: *if the null were true, how surprising would our observed $\hat\theta$ be?*

To answer that, we **standardize** $\hat\theta$ to form a test statistic:

$$z = \frac{\hat\theta - 0}{SE(\hat\theta)}$$

The logic runs as follows:

1. By the CLT, $\hat\theta$ is approximately Normal for large samples.
2. Under $H_0$, the mean of that distribution is 0.
3. Dividing a Normal random variable by its standard deviation gives a **standard Normal** $N(0,1)$. So under $H_0$, $z \;\dot\sim\; N(0,1)$.
4. We know the distribution of $z$ under the null — so we can compute the probability of observing a value as extreme as ours. That probability is the **p-value**.

A brief note on terminology: you may have heard this called a **t-test**. The classical t-test arises when data are drawn from a Normal distribution and the variance is unknown — the test statistic then follows an exact t-distribution, not a Normal. Our data are Bernoulli, not Normal, so there's no exact t-distribution result here. The CLT gives us approximate Normality, and what we're really doing is a **z-test**. In practice, for large samples the t-distribution and the standard Normal are nearly indistinguishable, so the distinction rarely matters — but it's worth understanding.

```{r hyp-test, echo=true}
theta_hat <- mean(draws_A) - mean(draws_B)
se        <- sqrt(pi_hat_A * (1 - pi_hat_A) / n +
                  pi_hat_B * (1 - pi_hat_B) / n)
z_stat    <- theta_hat / se
p_value   <- 2 * pnorm(-abs(z_stat))
```

```{r hyp-test-table, echo=false}
tibble(
  Statistic = c("Estimated difference (θ̂)", "Standard Error", "z-statistic", "p-value"),
  Value     = c(
    round(theta_hat, 4),
    round(se, 4),
    round(z_stat, 3),
    round(p_value, 4)
  )
) |>
  gt() |>
  tab_header(
    title    = "Two-Sample z-Test Results",
    subtitle = "H₀: θ = 0  vs.  H₁: θ ≠ 0"
  ) |>
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_options(table.width = pct(60))
```

With a p-value of `r round(p_value, 4)`, we have strong evidence against $H_0$. If the two CTAs were truly identical, there would be only about a `r round(p_value*100, 1)`% chance of observing a difference as large as ours (or larger) by chance alone. At the conventional 5% significance level, we reject $H_0$ and conclude that CTA A significantly outperforms CTA B.

---

## The T-Test as a Regression

Here's a surprising and useful fact: the two-sample test we just ran is **mathematically equivalent to a simple linear regression**.

To see this, we stack our data into a single dataset with two columns:

- $Y_i$: the outcome (1 = signed up, 0 = didn't)
- $D_i$: an indicator for which CTA the visitor saw (1 = CTA A, 0 = CTA B)

Then we fit the regression:

$$Y_i = \beta_0 + \beta_1 D_i + \varepsilon_i$$

The coefficients have clean interpretations:

- $\beta_0 = E[Y_i \mid D_i = 0] = \pi_B$ (the sign-up rate for CTA B)
- $\beta_0 + \beta_1 = E[Y_i \mid D_i = 1] = \pi_A$ (the sign-up rate for CTA A)
- Therefore, $\beta_1 = \pi_A - \pi_B = \theta$

The OLS estimate $\hat\beta_1$ is numerically identical to our $\hat\theta = \bar{X}_A - \bar{X}_B$.

```{r regression, echo=true}
df_reg <- tibble(
  Y = c(draws_A, draws_B),
  D = c(rep(1, n), rep(0, n))
)

fit <- lm(Y ~ D, data = df_reg)
```

```{r reg-table, echo=false}
tidy_fit <- broom::tidy(fit)

tidy_fit |>
  filter(term == "D") |>
  select(term, estimate, std.error, statistic, p.value) |>
  mutate(across(where(is.numeric), ~round(., 4))) |>
  gt() |>
  tab_header(
    title    = "OLS Regression: Sign-Up ~ CTA Indicator",
    subtitle = "Coefficient on D recovers the same estimate as the two-sample test"
  ) |>
  cols_label(
    term      = "Term",
    estimate  = "Estimate",
    std.error = "Std. Error",
    statistic = "t-statistic",
    p.value   = "p-value"
  ) |>
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_options(table.width = pct(80))
```

The regression coefficient $\hat\beta_1 =$ `r round(coef(fit)["D"], 4)` matches our earlier $\hat\theta$ exactly. The standard error and t-statistic are also very close (a minor difference arises because OLS assumes a pooled variance, while our earlier formula used separate variances per group — for large samples this distinction is negligible).

Why does this equivalence matter? Because the regression framework generalizes naturally. Want to control for user characteristics? Add covariates. Want to test three CTAs at once? Add more indicator variables. Want to check whether the CTA effect differs across device types? Add an interaction term. The t-test is a special case of regression, and understanding this connection unlocks a much more powerful toolkit.

---

## The Problem with Peeking

Your boss is impatient. She doesn't want to wait until all 1,000 visitors per group have been observed. Instead, she wants to run the hypothesis test **after every 100 visitors** — peeking at the results at $n = 100, 200, \ldots, 1000$ — and stop the experiment early if any peek shows a significant result.

This sounds reasonable. It's not.

The problem is a fundamental feature of classical hypothesis testing: each test at $\alpha = 0.05$ carries a **5% false positive rate** — a 5% chance of rejecting $H_0$ when it's actually true. When you run 10 such tests on accumulating data, you get 10 chances to make that mistake. The overall probability of at least one false rejection climbs well above 5%.

To see exactly how bad the situation is, we run a simulation in a world where **the null is true**: $\pi_A = \pi_B = 0.20$, so $\theta = 0$ and any "significant" result is a false positive.

```{r peeking-sim, echo=true}
set.seed(123)
n_experiments <- 10000
peeks         <- seq(100, 1000, by = 100)
false_positive <- logical(n_experiments)

for (i in 1:n_experiments) {
  A <- rbinom(1000, 1, 0.20)
  B <- rbinom(1000, 1, 0.20)

  significant <- FALSE
  for (peek in peeks) {
    a_sub  <- A[1:peek]
    b_sub  <- B[1:peek]
    p_a    <- mean(a_sub)
    p_b    <- mean(b_sub)
    se_p   <- sqrt(p_a * (1 - p_a) / peek + p_b * (1 - p_b) / peek)
    if (se_p == 0) next
    z      <- (p_a - p_b) / se_p
    pval   <- 2 * pnorm(-abs(z))
    if (pval < 0.05) { significant <- TRUE; break }
  }
  false_positive[i] <- significant
}

empirical_fpr <- mean(false_positive)
```

```{r peeking-table, echo=false}
tibble(
  Scenario              = c("Single test at n = 1,000", "Peeking at every 100 (10 tests)"),
  `False Positive Rate` = c("5.0% (by design)", paste0(round(empirical_fpr * 100, 1), "% (simulated)"))
) |>
  gt() |>
  tab_header(
    title    = "The Cost of Peeking",
    subtitle = "Empirical false positive rate under H₀: π_A = π_B = 0.20"
  ) |>
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style     = cell_fill(color = "#ffe0e0"),
    locations = cells_body(rows = 2)
  ) |>
  tab_options(table.width = pct(70))
```

When we peek 10 times and stop at the first significant result, our **empirical false positive rate is around `r round(empirical_fpr * 100, 1)`%** — far above the nominal 5%. In other words, even when there is truly no difference between the CTAs, we'd incorrectly declare a winner in roughly one out of every four or five experiments.

This is why **pre-registration matters** in A/B testing: you should decide your sample size *before* collecting data, run the test *once* at the end, and commit to that result. If you want the flexibility to stop early, you need more sophisticated methods (such as sequential testing or Bayesian approaches) that properly account for the multiple looks. Classical frequentist tests, applied naively with repeated peeking, will mislead you far more often than you think.
